\bf Title here

### 2020-09-17 paper/paper.Rmd
polish writing in Sec. 4.4 - 4.6

In projection pursuit, the projection bases $\mathbf{A}_{p \times d}$ are of dimension $p \times d$ with $p$ usually greater than two and hence can't be visualised in a 2D plot. An option to explore the searching space of these bases is to explore a reduced space via principal component analysis (PCA). The visualisation can thus be defined as

@@

Figure 5 shows the searching space plot of a geodesic search, colored by different searching stages. Starting from the start basis, a directional search is conducted in a narrow neighbourhood on five random directions. The best one is picked and an exact search finds the stepsize that maximises the index value and outputs the corresponding projection basis.

```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
dt <- bind_rows(holes_2d_better %>% mutate(algo = "not satisfactory"),
                holes_2d_better_max_tries %>% mutate(algo = "satisfactory")) %>%
 bind_rows(holes_2d_theory %>% mutate(algo = "satisfactory")) %>%
  bind_rows(holes_2d_theory %>% mutate(algo = "not satisfactory")) %>%
  mutate(algo = fct_relevel(algo, c("satisfactory", "not satisfactory")))%>%
  compute_pca() %>%
  `$`(aug)

dt %>% explore_proj_pca(facet = algo, size = 7, alpha = 0.5)
#dt %>% explore_proj_pca(facet = algo, axis = 2)
```

@@

all the plots above can be animated if wanted - importance is how the data is moving around

show an example of some snap shots  of the animation and provide a link... for animated pca plot

```{r animated-pca, fig.height = 8, fig.cap = "\\label{animated-pca} 16 frames from animated PCA plot ." }
rl <- lapply(list.files(here::here("figures/animated_pca/"),full.names = TRUE)[1:16], png::readPNG)
gl <-  lapply(rl, grid::rasterGrob)
gridExtra::grid.arrange(grobs=gl)
```

@@

show the path of the tour on a sphere. (full data dimension animation...)

include the tour animation (geozoo)

>>>>

Apart from checking the progression of an optimiser, another interesting aspect is to visualise how the search looks like in its parameter space. Given the orthonormality constraint, the projection bases $\mathbf{A}_{p \times d}$ lives on the surface of a $p \times d$ dimension sphere, where the dimension can easily go above five (5). Visualising the search paths on the original high diemnsional sphere would require skills from the viewer to preceive rotation of geometry in higher dimensional space (d > 3) while an easier alternative is to view the reduced space via some dimension reduction methods i.e. Principal component analysis. To better visual the serach path as an embedding of a hollow sphere, random points on the high dimensional sphere is generated using package `geozoo` and PCA is conducted on both the bases and the points on the surface of the sphere.

The visualisation can thus be defined as

@@

Figure \ref{toy-pca} shows the first two principal components of two search paths. Starting from the same position, two searching methods take different paths to get its final bases. There are two theoretical bases because they correspond to the same data projection $Y$ with a 180 degree rotation. What differentiate the two optimisers is that `search_better()` also involve random evaluation of points in the sphere during its sampling process while `search_geodesic()` doesn't and this stochastic rather than deterministic approach can be useful in complex scenario.

```{r toy-pca, fig.height = 6, fig.cap = "\\label{toy-pca}PCA plot of search geodesic colouring by info allows for better understanding of each stage in the geodesic search"}
path1 <- holes_1d_geo$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
path2 <- holes_1d_better$basis %>% 
  flatten_dbl() %>% 
  matrix(ncol = 5, byrow = TRUE)
set.seed(1)
sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
theoretical <- matrix(c(0, 1, 0, 0, 0,
                        0, -1, 0, 0, 0), nrow = 2, byrow = TRUE)

dt <- rbind(sphere, path1, path2, theoretical)
pca <- dt %>% prcomp()
#summary(pca)
pca_pred <- predict(pca) %>%
  as_tibble() %>%
  mutate(method = c(rep("sphere", nrow(sphere)),
                    holes_1d_geo$method,
                    holes_1d_better$method,
                    rep("theoretical", 2))) %>%
  group_by(method) %>%
  mutate(id = row_number(),
         id = ifelse(method == "sphere", 1, id)) %>%
  ungroup()

p <- pca_pred %>%
  ggplot(aes(x = PC1, y = PC2, color = method))+
  geom_point() +
  geom_point(data = pca_pred %>% filter(method == "theoretical"), 
             size = 10) + 
  geom_point(data = pca_pred %>% filter(method %in% c("search_better", "search_geodesic") & id == 1), size = 10) + 
  theme(aspect.ratio = 1) +
  scale_color_botanical(palette = "cherry")

p
```

@@

Animating the plots introduced above is useful, especially in the case of PCA polot since it shows the bases found by the optimiser in its natural order.

<!-- all the plots above can be animated if wanted - importance is how the data is moving around  -->

<!-- show an example of some snap shots  of the animation and provide a link... for animated pca plot -->

<!-- This is particularly useful when plotting the PCA plot since it allways us to tract the progress of the search. -->

<!-- The projection bases live in a high dimensional sphere, whch cna be simulated using `geozoo` package. -->

<!-- The reason to simulate random points on the hihg dimensiaon sphere is to emphasize  taht the projection bases lives on the surface of the sphere. -->

```{r cache = TRUE, eval=FALSE}
# rl <- lapply(list.files(here::here("figures/animated_pca_1d/"),full.names = TRUE)[62:78], png::readPNG)
# gl <-  lapply(rl, grid::rasterGrob)
# gridExtra::grid.arrange(grobs=gl)
ani <- p +
  transition_states(id) +
  shadow_mark()

animate(ani, nframes = 350)
anim_save(ani, nframes = 350, filename = "figures/animated_pca_1d")
```

@@

Viewing the bases on the reduced space via PCA shed some lights on the space the optimisers have explored, the visualisation on the original $p \times d$ dimension enables a stereoscopic view of the search. To view a high dimensional ($d \ge 3$) object on a screen, an approach is to play the rotation of the object in animation. This can be done via a regular grand tour or a slice tour [@laa2020slice], which emphasizes the points that are close to a section of the sphere.

Compare to the PCA plot, the animated rotation (tour) display gives a more holistic view of the search of the optimiser and these additional information gained from the animation is cruicial. This is because in essense, PCA presents one reduced space that maximises the variance and it inherently amplifies the spread of the randomly evaluated bases in the search since they have larger variance than the bases on the interpolation path. Hence when the interest is to view the interpolation path in the space, an animated rotation display provides the view of high dimension from different angles.

```{r eval = FALSE}
set.seed(1)
sphere <- geozoo::sphere.hollow(p = 5, n = 1000)$point
path1 <- holes_1d_geo$basis %>% 
  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
path2 <- holes_1d_better$basis %>% 
  flatten_dbl() %>% matrix(ncol = 5, byrow = TRUE)
dt <- rbind(sphere, path1, path2)
colnames(dt) <- c(map_chr(1:5, ~paste0("x", .x)))
pal <- c("#1B9E77","#D95F02","#7570B3", "#E7298A")
color <- c(rep(pal[1], nrow(sphere)),
           rep(pal[2], nrow(path1)),
           rep(pal[3], nrow(path2)))
animate_xy(dt, col = color, cex_slice = 0.8)
animate_slice(dt, col = color)

### 2020-09-17 paper/paper.tex
polish writing in Sec. 4.4 - 4.6

The projection bases \(\mathbf{A}_{p \times d}\) live on the surface of
a \(p \times d\) dimension sphere, where this dimension can easily go
from 5 to 10. This makes viewing the original space hard. An easier
alternative is to view the reduced space via some dimensional reduction
methods i.e.~Principal component analysis. The visualisation can thus be
defined as

@@

Figure \ref{toy-pca} shows the first two principal components of two
searching paths. A Hollow sphere with the same dimension is overlaid to
emphasize the imposition of orthonormal constraint. Starting from the
same position, two searching methods takes different paths to get to the
two final bases with the same data projection \(Y\) (the two correspond
to a rotation of the data in the projection). The distinction between
the two is that \texttt{search\_better()} also involve some random
evaluation in the sphere due to its sampling process while
\texttt{search\_geodesic()} doesn't. This stochastic rather than
deterministic approach can be useful in more complex scenario.

@@

An animated plot can be useful because it allows us to track the
progression of the optimsiation, especially in the case of inspecting
the search space.

@@

As mentioned previously, the orthonormal constraint regulates the
projection basis to be living on the surface of a unit sphere. The
visualisation of the projection basis on its original \(p \times d\)
dimension can be done through an animated rotation of the \(p \times d\)
diemnsion sphere with the searching path overlaid. The high dimensional
geometry can be simulated using the package \texttt{geozoo} and the
animated rotation can be done via a regular grand tour or a slice tour
\citep{laa2020slice}.

Compared with the PCA plot, this animated rotation display requires
skills of the viewer to preceive the high dimensional space in rotation
and what the viewers gain is a much holistic view of the path in the
space in the sense that PCA does a further projection on the projection
basis along with the hollow sphere, which gives the view that maixmises
the variance. While the animated rotation display gives the view from
different anlges and the PCA plot consitutes one frame from the animated
rotaiton display.

>>>>

Apart from checking the progression of an optimiser, another interesting
aspect is to visualise how the search looks like in its parameter space.
Given the orthonormality constraint, the projection bases
\(\mathbf{A}_{p \times d}\) lives on the surface of a \(p \times d\)
dimension sphere, where the dimension can easily go above five (5).
Visualising the search paths on the original high diemnsional sphere
would require skills from the viewer to preceive rotation of geometry in
higher dimensional space (d \textgreater{} 3) while an easier
alternative is to view the reduced space via some dimension reduction
methods i.e.~Principal component analysis. To better visual the serach
path as an embedding of a hollow sphere, random points on the high
dimensional sphere is generated using package \texttt{geozoo} and PCA is
conducted on both the bases and the points on the surface of the sphere.

The visualisation can thus be defined as

@@

Figure \ref{toy-pca} shows the first two principal components of two
search paths. Starting from the same position, two searching methods
take different paths to get its final bases. There are two theoretical
bases because they correspond to the same data projection \(Y\) with a
180 degree rotation. What differentiate the two optimisers is that
\texttt{search\_better()} also involve random evaluation of points in
the sphere during its sampling process while \texttt{search\_geodesic()}
doesn't and this stochastic rather than deterministic approach can be
useful in complex scenario.

@@

Animating the plots introduced above is useful, especially in the case
of PCA polot since it shows the bases found by the optimiser in its
natural order.

@@

Viewing the bases on the reduced space via PCA shed some lights on the
space the optimisers have explored, the visualisation on the original
\(p \times d\) dimension enables a stereoscopic view of the search. To
view a high dimensional (\(d \ge 3\)) object on a screen, an approach is
to play the rotation of the object in animation. This can be done via a
regular grand tour or a slice tour \citep{laa2020slice}, which
emphasizes the points that are close to a section of the sphere.

Compare to the PCA plot, the animated rotation (tour) display gives a
more holistic view of the search of the optimiser and these additional
information gained from the animation is cruicial. This is because in
essense, PCA presents one reduced space that maximises the variance and
it inherently amplifies the spread of the randomly evaluated bases in
the search since they have larger variance than the bases on the
interpolation path. Hence when the interest is to view the interpolation
path in the space, an animated rotation display provides the view of
high dimension from different angles.
